---
title: "MixVsSingleCompunds"
output:
    html_document:
        toc: true
        toc_float: true
        collapsed: true
        theme: 'simplex'
        highlight: 'espresso'
        code_folding: hide
---

### Data loading

```{r, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(SummarizedExperiment)  
  library(SEtools)
  library(edgeR)
  library(DT)
  library(pheatmap)
  library(plotly)
  library(dplyr)
  library(sva)
  library(ggrepel)
})
source("Functions/EDC_Functions.R")
source("Functions/CriFormatted.R")
load("Data/AllSEcorrected.RData",verbose=T)
load("Data/HormonalGenes/HormonalGenes.RData",verbose = T)
load("Data/DEAfetSingleCompunds.RData", verbose=TRUE)
load("Data/DEAorgSingleCompunds.RData", verbose=TRUE)
```

### MixN vs single compounds Fetal
#### Fetal MixN DEGs are specific for MixN treated samples
```{r, warning=FALSE, message=FALSE}
fetal <- SEs$chronic.fetal[,which(SEs$chronic.fetal$EXPO=="CNT" | SEs$chronic.fetal$EXPO=="1X" | SEs$chronic.fetal$EXPO=="1000X" | SEs$chronic.fetal$EXPO=="T3"| SEs$chronic.fetal$EXPO=="BPA0.04X")]

degs.fe <- row.names(DEAs$chronic.fetal)[which((abs(DEAs$chronic.fetal$logFC.EXPO1X)>0.5 | abs(DEAs$chronic.fetal$logFC.EXPO1000X)>0.5) & DEAs$chronic.fetal$FDR<=0.05 & DEAs$chronic.fetal$logCPM>0)]

sehm(fetal, assayName = "corrected", degs.fe,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"), main="MixNDEGs")
```

#### Fetal T3 treated samples respond to T3 known targets and the DEGs are specific for T3 samples

```{r, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5}
sehm(fetal[,which(SEs$chronic.fetal$EXPO=="CNT" | SEs$chronic.fetal$EXPO=="T3")], assayName = "corrected", Thyroid,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"),  main="T3 target genes")

degs.fetT3 <- row.names(res.fetT3)[which(abs(res.fetT3$logFC)> 0.5 & res.fetT3$FDR<=0.05 & res.fetT3$logCPM > 0)]
sehm(fetal, assayName = "corrected", degs.fetT3,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"), main="T3DEGs")


T3 <- res.fetT3
T3$genes <- row.names(T3)

MixN <- DEAs$chronic.fetal
MixN$genes <- row.names(MixN)
MixN$logFC <- (MixN$logFC.EXPO1X + MixN$logFC.EXPO1000X)/2

Comp <- compareResultsFCNew(T3, MixN, FDRth=0.05, FCth=2^0.5, FDRceil=1e-10, logCPMth=0, FCceil=5, 
                              title='Fetal: Comparing T3 to MixN (mean FC)', geneLabel=TRUE, topLab=-30)
Comp$Scatter

#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterFetal_MixN_T3.pdf', width=10, height=10)
#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterFetal_MixN_T3.png', width=10, height=10)
```

```{r}
# Check overlap between MixN and Thyroid
table(Comp$AllRes$Status)

## Fisher Test Significant
SigBoth <- dplyr::filter(Comp$Significant, Comp$Significant$Status=='SignificantA&B')

Contingency <- matrix(c(dim(dplyr::filter(SigBoth, log2FCA < 0 & log2FCB < 0 ))[1], 
                          dim(dplyr::filter(SigBoth, log2FCA >= 0 & log2FCB < 0 ))[1],
                           dim(dplyr::filter(SigBoth, log2FCA < 0 & log2FCB >= 0 ))[1], 
                            dim(dplyr::filter(SigBoth, log2FCA >= 0 & log2FCB >= 0 ))[1]),
                        nrow=2, dimnames=list(c('DownA', 'UpA'), c('DownB', 'UpB')))
Contingency
fisher.test(Contingency)
```


#### Fetal BPA treated samples show specific DEGs
```{r, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5}
degs.fetBPA <- row.names(res.fetBPA)[which(abs(res.fetBPA$logFC)> 0.5 & res.fetBPA$FDR<=0.05 & res.fetBPA$logCPM > 0)]

sehm(fetal, assayName = "corrected", degs.fetBPA,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"), main="BPADEGs")

BPA <- res.fetBPA
BPA$genes <- row.names(BPA)

MixN <- DEAs$chronic.fetal
MixN$genes <- row.names(MixN)
MixN$logFC <- MixN$logFC.EXPO1X

Comp <- compareResultsFCNew(BPA, MixN, FDRth=0.05, FCth=2^0.5, FDRceil=1e-10, logCPMth=0, FCceil=4, 
                              title='Fetal: Comparing BPA to MixN 1X', geneLabel=TRUE, topLab=-30)
Comp$Scatter

#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterFetal_MixN_BPA.pdf', width=10, height=10)
#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterFetal_MixN_BPA.png', width=10, height=10)
```

### MixN vs single compounds Organoids
#### Organoids MixN DEGs are specific for MixN treated samples
```{r, warning=FALSE, message=FALSE}
Organoids <- SEs$chronic.org[,which(SEs$chronic.org$EXPO=="CNT"|SEs$chronic.org$EXPO=="1X" | SEs$chronic.org$EXPO=="1000X" | SEs$chronic.org$EXPO=="T3"| SEs$chronic.org$EXPO=="BPA0.04X")]
degs.org <- row.names(DEAs$chronic.org)[which((abs(DEAs$chronic.org$logFC.EXPO1X)>0.5 | abs(DEAs$chronic.org$logFC.EXPO1000X)>0.5) & DEAs$chronic.org$FDR<=0.05 & DEAs$chronic.org$logCPM>0)]

sehm(Organoids, assayName = "corrected", degs.org,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"), main="MixNDEGs")
```

#### Organoids T3 treated samples respond to T3 known targets and the DEGs are specific for T3 samples
```{r, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5}
sehm(Organoids[,which(SEs$chronic.org$EXPO=="CNT" | SEs$chronic.org$EXPO=="T3")], assayName = "corrected", Thyroid,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"), main="T3 target genes")


degs.orgT3 <- row.names(res.orgT3)[which(abs(res.orgT3$logFC)> 0.5 & res.orgT3$FDR<=0.05 & res.orgT3$logCPM > 0)]

sehm(Organoids, assayName = "corrected", degs.orgT3,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"), main="T3DEGs")

T3 <- res.orgT3
T3$genes <- row.names(T3)

MixN <- DEAs$chronic.org
MixN$genes <- row.names(MixN)
MixN$logFC <- (MixN$logFC.EXPO1X + MixN$logFC.EXPO1000X)/2

Comp <- compareResultsFCNew(T3, MixN, FDRth=0.05, FCth=2^0.5, FDRceil=1e-10, logCPMth=0, FCceil=6, 
                              title='Comparing T3 to MixN (mean FC)', geneLabel=TRUE, topLab=-40)
Comp$Scatter

#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterOrg_MixN_T3.pdf', width=10, height=10)
#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterOrg_MixN_T3.png', width=10, height=10)
```

```{r}
# Check overlap between MixN and Thyroid
table(Comp$AllRes$Status)

## Contingency table on shared genes
SigBoth <- dplyr::filter(Comp$Significant, Comp$Significant$Status=='SignificantA&B')
ContingencyII <- matrix(c(dim(dplyr::filter(SigBoth, log2FCA < 0 & log2FCB < 0 ))[1], 
                          dim(dplyr::filter(SigBoth, log2FCA >= 0 & log2FCB < 0 ))[1],
                           dim(dplyr::filter(SigBoth, log2FCA < 0 & log2FCB >= 0 ))[1], 
                            dim(dplyr::filter(SigBoth, log2FCA >= 0 & log2FCB >= 0 ))[1]),
                        nrow=2, dimnames=list(c('DownA', 'UpA'), c('DownB', 'UpB')))
ContingencyII
prop.table(ContingencyII)
fisher.test(ContingencyII)

dplyr::filter(SigBoth, log2FCA < 0 & log2FCB > 0 ) %>% dplyr::pull(genes)
```

#### Organoids BPA treated samples show specific DEGs
```{r, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5}
degs.orgBPA <- row.names(res.orgBPA)[which(abs(res.orgBPA$logFC)> 0.5 & res.orgBPA$FDR<=0.05 & res.orgBPA$logCPM > 0)]

sehm(Organoids, assayName = "corrected", degs.orgBPA,  do.scale = T, show_rownames = F, anno_columns = c("EXPO2","MixtureBatch"),  main="BPADEGs")

BPA <- res.orgBPA
BPA$genes <- row.names(BPA)

MixN <- DEAs$chronic.org
MixN$genes <- row.names(MixN)
MixN$logFC <- MixN$logFC.EXPO1X

#Comp <- compareResultsFC(BPA, MixN, FDRth=0.05, FCth=2^0.5, FDRceil=1e-10, FCceil=8,logCPMth = 0, title='Comparing BPA to MixN 1X', geneLabel=TRUE)
#Comp$Scatter

Comp <- compareResultsFCNew(BPA, MixN, FDRth=0.05, FCth=2^0.5, FDRceil=1e-10, logCPMth=0, FCceil=4, 
                              title='Organoids: Comparing BPA to MixN 1X', geneLabel=TRUE, topLab=-30)
Comp$Scatter

#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterOrg_MixN_BPA.pdf', width=10, height=10)
#ggsave(Comp$Scatter, filename='Data/SingleCompounds/ScatterOrg_MixN_BPA.png', width=10, height=10)
```


```{r}
# Check overlap between MixN and BPA
table(Comp$AllRes$Status)

## Contingency table on shared genes
SigBoth <- dplyr::filter(Comp$Significant, Comp$Significant$Status=='SignificantA&B')
ContingencyII <- matrix(c(dim(dplyr::filter(SigBoth, log2FCA < 0 & log2FCB < 0 ))[1], 
                          dim(dplyr::filter(SigBoth, log2FCA >= 0 & log2FCB < 0 ))[1],
                           dim(dplyr::filter(SigBoth, log2FCA < 0 & log2FCB >= 0 ))[1], 
                            dim(dplyr::filter(SigBoth, log2FCA >= 0 & log2FCB >= 0 ))[1]),
                        nrow=2, dimnames=list(c('DownA', 'UpA'), c('DownB', 'UpB')))
ContingencyII
prop.table(ContingencyII)
fisher.test(ContingencyII)
```

### Tentative: overlap

N.B. Overlap analyses with the function below are not equivalent to the scatterplots, because here the selection of DEGs for MixN is done with an 'or' approach among the two concentrations for the FC (while above we use MixN 1x FC for BPA and the mean between the two concentrations for T3).

```{r}
library(overlapper)
m1 <- overlapper::multintersect(ll=list(fetalDEGs=degs.fe, fetalBPA=degs.fetBPA, fetalT3=degs.fetT3), universe=row.names(DEAs$chronic.fetal), two.tailed = F)

dotplot.multintersect(m1, sizeRange = c(0,15), th=0.05)
```


```{r}
m2 <- overlapper::multintersect(ll=list(orgDEGs=degs.org, organoidBPA=degs.orgBPA, organoidT3=degs.orgT3), universe=row.names(DEAs$chronic.org),two.tailed = F)

dotplot.multintersect(m2, sizeRange = c(0,15), th=0.05)
```



### Data praparation

For details on data filtering, normalization, batch correction and differential expression analysis, see [here](./DataPreparation/SVAcorrectedObjects.html) and [here](./DataPreparation/DEAsingleCompunds.html) 

***

```{r child='footer.Rmd'}
```


```{r SaveSession}
sessionInfo()
```

